# Reproducible Research: Peer Assessment 1
Miscellaneous setup
``` { r misc}
install.packages("xtable")
```

## Loading and preprocessing the data
The raw data is assumed to be int he working directory (this is a reasonable
assumption since the project istarted as a clone which contains the data).

Unzip and read the datafile.
```{r loadData}
fileName = 'activity'
if( !file.exists(paste0( fileName, '.csv'))) {
  unzip( paste0( fileName, '.zip'))
}
data = read.csv(paste0( fileName, '.csv'))
```
## Manipulate the Data
Remove NA data and change date strings to dat objects.
```{r showData}
data <-  na.omit(data) 
data$date <- as.Date(data$date, "%Y-%m-%d")
```
## What is mean total number of steps taken per day?
```{r plotData}
daily = tapply( data$steps, data$date, sum)
hist( daily, main = "Histogram of frequency of daily steps", xlab = "Number of Daily Steps")
```

Compute daily mean and median values.
```{r stats}
library(xtable)
dmean=tapply( data$steps, data$date, mean)
dmedian=tapply( data$steps, data$date, median)
dboth = cbind( dmean, dmedian)
colnames( dboth) <- c( "Mean", "Median" ) 
```
### Table of Daily Mean and Median Step Values
```{r, results='asis'}
print( xtable(dboth), type="html")
```

## What is the average daily activity pattern?
```{r plot}
imean = tapply( data$steps, data$interval, mean)
lmean = unlist(labels(imean))
plot( lmean,imean, type="l", xlab="Interval", ylab="Number of Steps", main="Average Daily Activity Pattern")
```
```{r vars}
maxSteps = max( data$steps)
maxInt = data$interval[ data$steps==max( data$steps)]
maxDate = data$date[ data$steps==max( data$steps)]
```
###Maximum Number of Steps
The maximum number of steps (`r maxSteps`) was in interval `r maxInt` on `r maxDate`.

## Imputing missing values
Replace NA values in number of steps with the daily average.
```{r missing}
data2 = read.csv(paste0( fileName, '.csv'))
data2$steps[is.na(data2$steps)]=tapply( data$steps, data$date, mean)
daily2 = tapply( data2$steps, data2$date, sum)
hist( daily2, main = "Histogram of frequency of daily steps (with imputed missing values)", xlab = "Number of Daily Steps")
```
```{r missingstats, results='asis'}
library(xtable)
dmean2=tapply( data2$steps, data2$date, mean)
dmedian2=tapply( data2$steps, data2$date, median)
dboth2 = cbind( dmean2, dmedian2)
colnames( dboth2) <- c( "Mean", "Median" ) 
```
### Table of Daily Mean and Median Step Values with Imputed Missing Values
```{r, results='asis'}
print( xtable(dboth2), type="html")
```
## Are there differences in activity patterns between weekdays and weekends?
Yes, the data shows a difference between weekday and weekend activity.  
```{r weekdays}
data$isWeekday <- factor(weekdays(data$date) %in% c('Saturday','Sunday'), labels=c("weekday", "weekend"))
```
```{r panelplot}
library(lattice)
weekdays <- subset(data,data$isWeekday == "weekday")
weekends <- subset(data,data$isWeekday == "weekend")
xyplot( data$steps~data$interval | data$isWeekday, ylab="Number of steps", xlab="Interval", type="l", scales=list( alternating=TRUE),layout=c(1,2))
```